---
title: "LP optimize"
author: "Dr. J"
date: "3/15/2020"
output: html_document
---

# Data Analytics: optimization  
Optimization is an important part of data analytics.  In this problem we solve a linear programming example:

Maximize  
$Profit = 5 \times x_1 + 7 \times x_2$  

subject to the constraints:  
$x_1 \le 16$  
$2 \times x_1 + 3 \times x_2 \le 19$  
$x_1 + x_2 \le 8$  and  
$x_1 , x_2 \ge 0$  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


 Import lpSolve package
```{r}
library(lpSolve)
```


Set coefficients of the objective function
```{r}
f.obj <- c(5, 7)
```


Set matrix corresponding to coefficients of constraints by rows
Notice that the non-negative constraints on $x_1$ and $x_2$ are automatically assumed  

```{r}
f.con <- matrix(c(1, 0,
                  2, 3,
                  1, 1), nrow = 3, byrow = TRUE)

```


Set inequality signs:  
```{r}
f.dir <- c("<=",
           "<=",
           "<=")
```

Set right hand side coefficients
```{r}
f.rhs <- c(16,
           19,
           8)
```


Find maximum profit
```{r}
maxP = lp("max", f.obj, f.con, f.dir, f.rhs)
print(maxP)
```


print the values of $x_1$ and $x_2$ that maximize profit:
  
```{r}
print(paste('x1: ',maxP$solution[1], '  x2: ',maxP$solution[2]))
```


## Sensitivities
```{r}
lp("max", f.obj, f.con, f.dir, f.rhs, compute.sens=TRUE)$sens.coef.from
lp("max", f.obj, f.con, f.dir, f.rhs, compute.sens=TRUE)$sens.coef.to
```



 Dual Values (first dual of the constraints and then dual of the variables)  
 Duals of the constraints and variables are mixed  
 
```{r}
lp("max", f.obj, f.con, f.dir, f.rhs, compute.sens=TRUE)$duals
```



Duals lower and upper limits

```{r}
lp("max", f.obj, f.con, f.dir, f.rhs, compute.sens=TRUE)$duals.from
lp("max", f.obj, f.con, f.dir, f.rhs, compute.sens=TRUE)$duals.to
```

